---
title: "Basic of Selecting Data"
author: "Sandesh"
date: "2021-11-07"
output: html_document
---

In this article we will discuss the basic of the data manipulation --
slicing and selecting. As we have seen, majority of time, we need to work
with a subset of the data for the analysis. We might need just need to
work on the description on one variable. Such situation is common in the
data analytics workflow. So, selection and slicing of right data in
efficient manner is an essential skill for a data analyst. The article
will discuss about selecting data for following data structures:

1.  Vector

2.  List

3.  Data Frame

**Vectors:**

Vector is the basic data-structure in R. The element of the vectors must
be of same data types. Meaning a vector can't contain the mixed
variables like integer, string, Boolean etc. Vectors are one
dimensional.

For the selection of a specific element of the vector, we simply put the
index number inside the big bracket. For instance, if we have to select
the 2nd element of vector X, our code will be X[2]. Let me remind you
one more time, unlike other majority of programming languages, the
indexing of R always start at 1. We will explore the details in the demo
below.

```{r Vectors Selection}

#Making some vectors

vec1 <- c(11,23,34,46,51)
vec2 <- c('this','is','a','vector')
vec3 <- c(TRUE, FALSE, FALSE, TRUE)

#Selecting the single values
vec1[2]

#selecting the multiple values
vec1[c(2,3)] #in this example we are selecting the 2nd and 3rd element of vec1

#selecting after a certain index
vec1[1:5] #we selected everything from index 1 to 5.
vec2[-1]  # -1 will include everything except the first element
```

**List**

List are different than the vector because they can have multiple data
types.The list are formulated by using list() instead of c() in vectors.
List sometimes can be recursive as a list can contain list within the
list. I know that's confusing but will explain on the demo below:

```{r List Selection}
lst1 <- list(100, 101, 102, 103, 104, 105)
lst2 <- list(155, 'program', TRUE, 3.25, 200)
lst3 <- list(lst1, lst2 , 534, 546, TRUE, 1.98) #the complicated list

#selection in the plain list
lst1[2] #this is similar to the vector
lst2[3:5]

#selection in the complicated one
lst3[[2]][2]  
#first double bracket is selecting the 2nd element of the lst3 
#which is lst2, then we are selecting the 2nd element of lst2.

lst3[[1]][2:5] 
#another similar example like above but we selecting from all 
#elements between 2nd and 5th. 

```

**Data Frame**

Data Frame are the general structure that we usually see in our analysis
project. Think them as the Excel sheet- with the rows and columns. The proper 
methods of selecting right data from the data frame plays a vital role in data
analysis. 

```{r DataFrame}
data= iris #loading the iris data and saving in the 'data' variable
head(data) #looking at the first few rows of the dataset.


#QUESTION 1: How do you select the value in 3rd column and 2nd row?
row_number = 2
column_number =3

data[row_number,column_number] #1.4 is our output. 


#QUESTION 2: How to select all the rows of the 2nd column?
data$Sepal.Width #the '$' sign in the easiest way to select a column.


#QUESTION 3:How to select multiple column?
data[1:10,c(2,3)] #by indexing. Making vector of needed column.
data[1:10,c('Sepal.Width','Sepal.Length')] #by using column name


#QUESTION 4: How to select first 10 rows of 2nd and 3rd column
data[1:10,c(2,3)]


#QUESTION 5: Select all the 'Sepal.Length' that is greater than 7
data[data$Sepal.Length>7,] 
#in the row we defined our condition and in column we left empty because we 
#needed all the columns

```

In this article, I just showed the basic of selecting data in R. Indeed,
these method can be made more complex to subset various data. Similar to
'QUESTION 5', we can use the filter conditions to select different variety of 
data. However, this article is intended for the people who are getting started 
with R. I will post much advance methods in future posts.

Thank You